---
title: "Geocoder Services"
output: rmarkdown::html_vignette
description: >
  Documentation on the supported geocoder services and their query parameters and settings
vignette: >
  %\VignetteIndexEntry{Geocoder Services}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
set.seed(42)

library(DT)
library(tidygeocoder)
library(gt)
library(dplyr)
```

## Overview

The supported geocoder services are shown in the table below. The `method` is used to select the geocoder service in tidygeocoder functions such as `geo()` and `reverse_geo()`. The usage rate limitations are listed for the free tier of the service when applicable and many services have faster rates available with paid plans.

Also note that there are many other considerations when selecting a geocoder service such as if the service uses open source data with permissive licensing or if there are restrictions on how you can use the data provided by the service. Refer to each service's documentation for details.


```{r, echo = FALSE, message = FALSE, warning = FALSE, output = 'asis'}
library(dplyr)
check_mark <- "\U2705" #unicode character for heavy white check mark

geocoder_summary_table <-
  tidygeocoder::api_info_reference %>%
    mutate(
      service = paste0(
        '[', method_display_name, '](', site_url, ')'
      ),
      batch_geocoding = ifelse(method %in% names(tidygeocoder:::batch_func_map), check_mark, ''),
      api_key_required = ifelse(method %in% tidygeocoder::api_key_reference[['method']], check_mark, ''),
      api_documentation = paste0(
        '[docs](', api_documentation_url, ')'
      )
    ) %>%
    left_join(tidygeocoder::min_time_reference %>% select(method, description), by = 'method') %>%
    select(service, method, api_key_required, batch_geocoding, usage_limitations = description, api_documentation) %>%
    mutate(across(method, function(x) stringr::str_c('`', x, '`'))) %>% # format method column
    tidyr::replace_na(list(usage_limitations = ''))
  
# Format column names
colnames(geocoder_summary_table) <- colnames(geocoder_summary_table) %>%
  stringr::str_replace_all('_', ' ') %>%
  stringr::str_to_title() %>%
  stringr::str_replace_all('Api', 'API')

geocoder_summary_table %>%
  knitr::kable()
```

**Highlights:**

-   The US Census service does not support reverse geocoding and is limited to the United States.
-   Geocodio is limited to the United States and Canada.
-   The Census and Nominatim ("osm") services are free and do not require an API key.
-   ArcGIS can be used for free without an API key, but some features require a paid subscription and API key.
-   Geoapify, Geocodio, Location IQ, OpenCage, Mapbox, HERE, TomTom, MapQuest, and Bing are commercial services that offer free tiers.
-   The Google service has no free tier and [bills per query](https://developers.google.com/maps/documentation/geocoding/usage-and-billing).

**Other Notes:**

-   The US Census service supports street-level addresses only (ie. "11 Wall St New York, NY" is OK but "New York, NY" is not).
-   The Mapbox service is capable of performing batch geocoding when using the [permanent endpoint](https://docs.mapbox.com/api/search/geocoding/#batch-geocoding), but this capability is not currently implemented in tidygeocoder. If you'd like to add this capability to the package see [issue \#73](https://github.com/jessecambon/tidygeocoder/issues/73).
-   The ArcGIS service is capable of performing [batch geocoding](https://developers.arcgis.com/rest/geocode/api-reference/geocoding-geocode-addresses.htm) but this capability is not currently implemented in tidygeocoder. If you'd like to add this capability see [#102](https://github.com/jessecambon/tidygeocoder/issues/102).
-   For the ArcGIS service, an API Key is not strictly required if the service is used for search capabilities only (see [Free vs. paid operations](https://developers.arcgis.com/rest/geocode/api-reference/geocoding-free-vs-paid.htm)). It is possible to include an API Key on the request via the `custom_query` parameter:

``` r
tidygeocoder::geo(address = "New York, USA", method = "arcgis",
  custom_query = list(token = "<API_KEY>"))
```

## Usage Notes

- When used in batch mode, the **US Census** geocoder will return NA data when there are multiple results available for an address. The expectation is that you would see that a "Tie" is indicated and use single address geocoding to return the results for these addresses. See [#87](https://github.com/jessecambon/tidygeocoder/issues/87) for details.
- When performing reverse geocoding, the **Mapbox** service requires a `types` parameter to be set if `limit > 1`. See [#104](https://github.com/jessecambon/tidygeocoder/issues/104).
- The **Bing** batch geocoder does not use the `limit` parameter ([#106](https://github.com/jessecambon/tidygeocoder/issues/106)).

## API Parameters

The `api_parameter_reference` maps the API parameters for each geocoder service to a common set of "generic" parameters. The `generic_name` below is the generic parameter name while the `api_name` is the parameter name for the specified geocoder service (`method`). Refer to `?api_parameter_reference` for more details.

```{r, echo = FALSE}
api_parameter_reference %>% 
  mutate(across(c(method, generic_name, api_name), as.factor)) %>%
  datatable(filter = 'top', rownames = FALSE, 
  options = list(
    lengthMenu = c(5, 10, 15, 20, nrow(.)),
    pageLength = 10,
    autoWidth = TRUE)
  )
```

## API Key Retrieval

API keys are retrieved from environmental variables. The name of the environmental variable used for each service is stored in the `api_key_reference` dataset. See `?api_key_reference`.

```{r, echo = FALSE}
api_key_reference %>%
  gt() %>%
  opt_table_outline() %>%
  opt_table_lines() %>%
  tab_options(column_labels.font.weight = 'bold')
```

## Minimum Time Per Query

The minimum time (in seconds) required per query to comply with the usage limitations policies of each geocoder service is stored in the `min_time_reference` dataset. See `?min_time_reference`.

```{r, echo = FALSE}
min_time_reference %>%
  gt() %>%
  opt_table_outline() %>%
  opt_table_lines() %>%
  tab_options(column_labels.font.weight = 'bold')
```

<!-- https://bookdown.org/yihui/rmarkdown-cookbook/results-asis.html -->
Links to the usage policies for each geocoder service:

```{r, echo = FALSE, results = 'asis'}
cat(tidygeocoder:::get_api_usage_bullets(), sep = '\n')
```

## Batch Query Size Limits

The maximum number of inputs (geographic coordinates or addresses) per batch query for each geocoder service is stored in the `batch_limit_reference` dataset. See `?batch_limit_reference`.

```{r, echo = FALSE}
batch_limit_reference %>%
  gt() %>%
  fmt_number(columns = 'batch_limit', decimals = 0) %>%
  opt_table_outline() %>%
  opt_table_lines() %>%
  tab_options(column_labels.font.weight = 'bold')
```