---
title: "Introduction to tidygeocoder"
output: rmarkdown::html_vignette
description: >
  Start here if this is your first time using tidygeocder.
vignette: >
  %\VignetteIndexEntry{Introduction to tidygeocoder}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
set.seed(42)
```

Geocoder services are used to provide data (such as latitude and longitude coordinates) about locations (**forward geocoding**) or to provide data about the location of specific latitude and longitude coordinates (**reverse geocoding**). The goal of tidygeocoder is to make working with geocoder services easy. 

**For forward geocoding**, you can use either the `geocode()` or `geo()` functions which accept dataframe and character inputs respectively. The `geocode()` function extracts address columns from the input dataframe and passes them to `geo()`. All extra arguments (`...`) given to `geocode()` are passed to `geo()` so refer to the documentation in `geo()` for all the possible arguments you can give to the `geocode()` function.

**For reverse geocoding**, you can use the `reverse_geocode()` function which takes a dataframe input or `reverse_geo()` if you want to pass latitude and longitude values directly as vectors. The `reverse_geocode()` function extracts latitude and longitude values from the input dataframe and passes them to `reverse_geo()`. All extra arguments (`...`) given to `reverse_geocode()` are passed to `reverse_geo()` so refer to the documentation in `reverse_geo()` for all the possible arguments you can give to the `reverse_geocode()` function.

## Forward Geocoding

```{r setup, warning = FALSE, message = FALSE}
library(tibble)
library(DT)
library(dplyr)
library(tidygeocoder)

address_single <- tibble(singlelineaddress = c('11 Wall St, NY, NY', 
                    '600 Peachtree Street NE, Atlanta, Georgia'))
address_components <- tribble(
  ~street                      , ~cty,               ~st,
  '11 Wall St',                  'NY',               'NY',
  '600 Peachtree Street NE',     'Atlanta',          'GA'
)
```

You can use the `address` argument to specify single-line addresses. Note that when multiple addresses are provided, the batch geocoding functionality of the Census geocoder service is used. Additionally, `verbose = TRUE` displays logs to the console.

```{r}
address_single %>% geocode(address = singlelineaddress, method = 'census',
                           verbose = TRUE)
```

Alternatively you can run the same query with the `geo()` function by passing the address values from the dataframe directly. In either `geo()` or `geocode()`, the `lat` and `long` arguments are used to name the resulting latitude and longitude fields. Here the `method` argument is used to specify the OSM (Nominatim) geocoder service. Refer to the `geo()` function documentation for the possible values of the `method` argument.

```{r}
geo(address = address_single$singlelineaddress, method = 'osm', 
    lat = latitude, long = longitude)
```

Instead of single-line addresses, you can use any combination of the following arguments to specify your addresses: `street`, `city`, `state`, `county`, `postalcode`, and `country`. 

```{r}
address_components %>% geocode(street = street, city = cty, state = st,
                               method = 'census')
```

The `cascade` method first tries to use one geocoder service and then again attempts to geocode addresses that were not found using a second geocoder service. By default it first uses the Census Geocoder and then OSM, but you can specify any two methods you want (in order) with the `cascade_order` argument.

```{r}
addr_comp1 <- address_components %>% 
  bind_rows(tibble(cty = c('Toronto', 'Tokyo'), country = c('Canada', 'Japan')))

addr_comp1 %>% geocode(street = street, state = st, city = cty,
                       country = country, method = 'cascade')
```

To return more data than just the latitude and longitude coordinates, specify `full_results = TRUE`. Additionally, for the Census geocoder you can get fields for geographies such as Census tracts by specifying `return_type = 'geographies'`. Be sure to use `full_results = TRUE` with `return_type = 'geographies'` in order to allow the Census geography columns to be returned.

```{r}
census_full1 <- address_single %>% geocode(address = singlelineaddress, 
      method = 'census', full_results = TRUE, return_type = 'geographies')
```

```{r, echo = FALSE}
knitr::kable(census_full1)
```

As mentioned earlier, the `geocode()` function passes addresses in dataframes to the `geo()` function for geocoding so we can also directly use the `geo()` function in a similar way:

```{r}
salz <- geo('Salzburg, Austria', method = 'osm', full_results = TRUE)
```

```{r, echo = FALSE}
knitr::kable(salz)
```

## Reverse Geocoding

Reverse geocoding is structured in a very similar way to forward geocoding and many of the parameters that are used are the same. Instead of `geocode()` you'll use `reverse_geocode()` and instead of `geo()` you'll use `reverse_geo()`. Latitude and longitudes are provided as inputs and the address is returned. 

We can use `reverse_geocode()` to geocode latitude and longitude coordinates contained in a data frame. The `lat` and `long` arguments specify the columns that contain the latitude and longitude data. The `address` argument can be used to specify the single line address column name that is returned from the geocoder. The `

Just as with forward geocoding, the `method` argument is used to specify the geocoder and 

```{r}
lat_longs1 <- tibble(latitude = c(38.895865, 43.6534817), 
                longitude = c(-77.0307713, -79.3839347))

rev1 <- lat_longs1 %>%
  reverse_geocode(lat = latitude, long = longitude, address = addr,
                  method = 'osm', verbose = TRUE)
```

```{r, echo = FALSE}
knitr::kable(rev1)
```

The same query can also be performed by passing the latitude and longitudes directly to the `reverse_geo()` function. Here we will use `full_results = TRUE`.

```{r}
 rev2 <- reverse_geo(lat = lat_longs1$latitude, long = lat_longs1$longitude,
                  method = 'osm', full_results = TRUE)
glimpse(rev2)
```


## Working With Messy Data

Only unique data (either addresses or latitudes and longitudes) is passed to geocoder services even if your data contains duplicates. Missing/NA and blank inputs are excluded from queries. Also latitudes and longitudes are limited to possible values.

```{r}
duplicate_addrs <- address_single %>%
  bind_rows(address_single) %>%
  bind_rows(tibble(singlelineaddress = rep(NA, 3)))

duplicates_geocoded <- duplicate_addrs %>%
  geocode(singlelineaddress, verbose = TRUE)
```

```{r, echo = FALSE}
knitr::kable(duplicates_geocoded)
```

As shown above, duplicates will not be removed from your results by default. However, you can return only unique results by using `unique_only = TRUE`. Note that passing `unique_only = TRUE` to `geocode()` will result in the original dataframe format (including column names) to be discarded in favor of the standard field names (ie. "address", "city", "state", etc.).

```{r}
duplicate_addrs %>%
  geocode(singlelineaddress, unique_only = TRUE)
```

## Advanced Usage

The `limit` argument can be specified to return multiple matches per address if available:

```{r}
geo_limit <- geo(c('Lima, Peru', 'Cairo, Egypt'), method = 'osm', 
    limit = 3, full_results = TRUE)

glimpse(geo_limit)
```

To directly specify specific API parameters for a given `method` you can use the `custom_query` parameter. For example, [the Nominatim (OSM) geocoder has a 'polygon_geojson' argument](https://nominatim.org/release-docs/develop/api/Details/#parameters) that can be used to return GeoJSON geometry content. To pass this parameter you can insert it with a named list using the `custom_query` argument:

```{r}
cairo_geo <- geo('Cairo, Egypt', method = 'osm', full_results = TRUE,
    custom_query = list(polygon_geojson = 1), verbose = TRUE)

glimpse(cairo_geo)
```

To test a query without sending any data to a geocoder service, you can use `no_query = TRUE` (NA results are returned).

```{r}
geo(c('Vancouver, Canada', 'Las Vegas, NV'), no_query = TRUE, 
    method = 'osm')
```

Here are some additional usage notes for the `geocode()`, `geo()`, `reverse_geocode()`, and `reverse_geo()` functions:

* To specify a custom geocoder API URL, use the `api_url` argument. Alternatively, the `iq_region` and `geocodio_v` arguments are helper functions for customizing the API URL. 
* If not specified manually, the `min_time` argument will default to a value based on the maximum query rate of the given geocoder service. If you are using a local Nominatim server or have a commercial geocoder plan that has less restrictive usage limits then you can manually set `min_time` to a lower valuer (such as 0).
* By default batch geocoding will be used if available when more than one address is provided, otherwise single address geocoding will be used. To override this default, use the `mode` argument.

## API Reference

You can refer to the `api_parameter_reference` dataset to see which which parameters are supported with each geocoder service. This dataset is displayed below. 

Refer to `?api_parameter_reference` for more details and links to the API documentation for each geocoder service.

```{r, echo = FALSE}
api_parameter_reference %>% 
#  select(-required) %>%
  mutate(across(c(method, generic_name, api_name), as.factor)) %>%
  datatable(filter = 'top', rownames = FALSE, 
  options = list(lengthMenu = c(5, 8, 12, nrow(.)), pageLength = 12, autoWidth = TRUE))
```

