---
title: "Introduction to tidygeocoder"
output: rmarkdown::html_vignette
description: >
  Start here if this is your first time using tidygeocder.
vignette: >
  %\VignetteIndexEntry{Introduction to tidygeocoder}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
set.seed(42)
```

Geocoding services are used to provide data about addresses such as longitude and latitude coordinates. The goal of tidygeocoder is to make getting data from these services easy. The two main functions to use are `geocode()` which takes a dataframe as an input and `geo()` which takes character values or vectors of character values.

Before you jump into using tidygeocoder, you'll need to make a couple of decisions:

1. **Which geocoding service do you want to use?** Specify with the `method` argument:
  * `'census'` : the US Census geocoder service. Covers street level addresses for the US.
  * `'geocodio'` : a commercial geocoder service for the US and Canada.
  * `'osm'` : the Nominatim service queries OSM (Open Street Map) data. 
  * `'iq'` : a commercial offering of the Nominatim (OSM) service
  * `'cascade'`: this method first calls one geocoder service and if no results are found calls another geocoder service.
  
2. **Do you want to use single-line addressees or address components?** (ie. `address = '60 Main St, Pawnee, IN'` compared to `street = '60 Main St', city = 'Pawnee', state = 'IN'`)
  * For single-line addresses, use the `address` argument
  * Alternatively, you can use some combination of the `street`, `city`, `state`, `county`, `postalcode`, and `country` arguments. Note that not all of these parameters are required and geocoder services don't necessarily support all of these inputs, refer to the API documentation of the geocoder services for details. For example, the Census geocoder is only for the United States and has no reason to support a `country` input. 

## Making a Query

```{r, warning = FALSE, message = FALSE}
library(tibble)
library(dplyr)
library(tidygeocoder)

address_single <- tibble(singlelineaddress = c('11 Wall St, NY, NY', '600 Peachtree Street NE, Atlanta, Georgia'))
address_components <- tribble(
  ~str                          , ~cty,             ~st,
  '11 Wall St',                  'NY',               'NY',
  '600 Peachtree Street NE',     'Atlanta',          'GA'
)
```

The `geocode()` function takes a dataframe input with address data. Note that all parameters are passed to the `geo()` function so refer to that function's documentation for more details on possible arguments.

Use the `address` argument for single line addresses:

```{r}
address_single %>% geocode(address = singlelineaddress, method = 'census')
```

Alternatively use any combination of the following arguments to specify addresses: `street`, `city`, `state`, `county`, `postalcode`, and `country`.

```{r}
address_components %>% geocode(street = str, city = cty, state = st, method = 'census')
```

To return data other than latitude and longitude, specify `full_results = TRUE`. Additionally, for the census geocoder you can get details on  geographies such as census tracts by specifying `return_type = 'geographies'`.

```{r}
census_full1 <- address_single %>% geocode(address = singlelineaddress, 
      method = 'census', full_results = TRUE, return_type = 'geographies')
glimpse(census_full1)
```

## The cascade method

The `cascade` method by default first tries to use the census geocoder and then if no results are found then it attempts to use the OSM (Nominatim) geocoder. However, you can specify any two methods you would like to try in the order you'd like to try them using the `cascade_order` argument. 

```{r}
addr_comp1 <- address_components %>% 
  bind_rows(tibble(cty=c('Toronto', 'Tokyo'), country = c('Canada', 'Japan')))

addr_comp1 %>% geocode(street = str, state = st, city = cty, country = country, method = 'cascade', verbose = TRUE)
```


## The geo function

As an alternative to using the `geocode()` function, which operates on dataframes, you can pass address values directly to the `geo()` function as character values.

```{r}
salz <- geo('Salzburg, Austria', method = 'osm', full_results = T, verbose = T)
glimpse(salz)
```

Here is how you would geocode an address using the Census geocoder:

```{r}
geo(address = "600 Peachtree Street NE, Atlanta, Georgia", method = "census")
```

Alternatively, you can specify an address in components:

```{r}
geo(street = "600 Peachtree Street NE", city = "Atlanta", state = "Georgia", method = "census")
```

To use a different geocoder service, you can specify it with the `method` argument. To geocode multiple addresses, pass them as a vector. In this case we are using the Nominatim (OSM) geocoder service.

```{r}
geo(address = c("Tokyo, Japan", "Lima, Peru", "Nairobi, Kenya"), method = 'osm')
```
